Test susieI using whole genome, s40.22 samples

### 2020-11-13 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
pip cali

```{r snp}
total <- 1e3
n <- 22542
p1 <- pow(total, n, 0.0276, 5e-8)
print(p1)
```

@@

```{r gene1}
p2 <- pow(total, n, 0.025, 1e-5)
print(p2)
```

@@

```{r gene2}
p3 <- pow(total, n, 0.045, 1e-5)
print(p3)
# save(p1,p2,p3, file = "data/power_s40.22.Rd")
```

@@

cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5,]))
cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1 ,]))

@@

@@

## susieI
* Regions: all regions, 500kb uniform regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param1susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

@@

@@

@@

>>>>

```{r snp}
load("data/power_s40.22.Rd")
total <- 1e3
n <- 22542
#p1 <- pow(total, n, 0.0276, 5e-8)
print(p1)
```

@@

```{r gene1}
#p2 <- pow(total, n, 0.025, 1e-5)
print(p2)
```

@@

```{r gene2}
#p3 <- pow(total, n, 0.045, 1e-5)
print(p3)
#save(p1,p2,p3, file = "data/power_s40.22.Rd")
```

@@

cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5, ]))
cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1, ]))

@@

# TWAS FDP

```{r PIPplot functions}

library(ggplot2)
library(cowplot)
library(plotrix)

.exob <- function(pips, ifcausal, mode = c("PIP", "FDR")){
  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
  if (mode == "PIP") {
    ex = c(by(pips, a_bin, FUN = mean))
    ob = c(by(ifcausal, a_bin, FUN = mean))
  } else if (mode == "FDR"){
    ex = c(by(pips, a_bin, FUN = mean))
    ob = 1 - c(by(ifcausal, a_bin, FUN = mean))
  }
  return(list("expected" = ex, "observed" = ob))
}

dot_plot = function(df, main) {
        ggplot(df, aes(x=mean_pip, y=observed_freq)) + 
          geom_errorbar(aes(ymin=observed_freq-se, ymax=observed_freq+se), colour="black", size = 0.5, width=.01) +
          geom_point(size=1.5, shape=21, fill="#002b36") + # 21 is filled circle
          xlab("Expected") +
          ylab("Observed") +
          coord_cartesian(ylim=c(0,1), xlim=c(0,1)) +
          geom_abline(slope=1,intercept=0,colour='red', size=0.2) +
          ggtitle(main) +
          expand_limits(y=0) +                        # Expand y range
          theme_cowplot()}

cp_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){

  # ifcausal:0,1, runtag: for adding std.
  if (is.null(runtag)){
     se = 0 
  } else{
    dflist <- list()
    for (rt in unique(runtag)){
      pips.rt <- pips[runtag == rt]
      ifcausal.rt <- ifcausal[runtag == rt]
      dflist[[rt]] <- .exob(pips.rt, ifcausal.rt, mode = mode)
    }
    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")))
    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")))
    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error)
  }

  df <- data.frame("mean_pip" = mean_pip, "observed_freq"= observed_freq, "se" = se)
  dot_plot(df, main = main)

  #plot(Expected, Observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main, ...)
  #lines(x = c(0,1), y = c(0,1), col ="grey", lty = 2)
}

caliPIP_plot <- function(pipfs, format = "susie", main = "PIP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    res <- fread(pipfs[i], header = T)
    res <- data.frame(res[res$type  =="gene", ])
    res$runtag <- i
    if (format == "susie"){
       res <- rename(res, c("susie_pip" = "pip"))
    } else if (format == "SER"){
       res <- rename(res, c("SERpip" = "pip"))
    }
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  fig <- cp_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
  return(fig)
}

# pipfs is just for ifcausal info
caliFDP_plot <- function(gwasfs, pipfs,  format= "susie",  main = "FDP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    pipres <- fread(pipfs[i], header = T)
    pipres <- data.frame(pipres[pipres$type  =="gene", ])
    pipres$runtag <- i
    if (format == "susie"){
       pipres <- rename(pipres, c("susie_pip" = "pip"))
    } else if (format == "SER"){
       pipres <- rename(pipres, c("SERpip" = "pip"))
    }
    gwasres <- read.table(gwasfs[i], header = T)
    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
    gwasres$FDR <- p.adjust(gwasres$PVALUE, method = "fdr")
    res <- merge(gwasres, pipres, by = "name", all = T)
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }
  fig <- cp_plot(df$FDR, df$ifcausal, df$runtag, mode ="FDR", main = main)
  cat("FDP at bonferroni corrected p = 0.05: ", 1 - mean(df[df$PVALUE < 0.05 /dim(df)[1], "ifcausal"]))
  return(fig)
}

scatter_plot_PIP_p <- function(pipfs, gwasfs, pipformat = "susie", main ="PIP-p"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    pipres <- fread(pipfs[i], header = T)
    pipres <- data.frame(pipres[pipres$type  =="gene", ])
    pipres$runtag <- i
    if (pipformat == "susie"){
       pipres <- rename(pipres, c("susie_pip" = "pip"))
    } else if (pipformat == "SER"){
       pipres <- rename(pipres, c("SERpip" = "pip"))
    }
    gwasres <- read.table(gwasfs[i], header = T)
    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
    res <- merge(gwasres, pipres, by = "name", all = T)
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  df <- rename(df, c( "PVALUE" = "TWAS.p"))
  df[,"TWAS.p"] <- -log10(df[, "TWAS.p"])

  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))

  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))

  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
  #                                                                  "\nChr: ", chr,  "\nPos:", pos))

}

```

```{r TWAS FDP, fig.height=4, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

f1 <- caliFDP_plot(gwasfs[1:5], susieIfs2[1:5], format= "susie", main = "TWAS FDP (low power)")
f2 <- caliFDP_plot(gwasfs[6:10], susieIfs2[6:10], format= "susie", main ="TWAS FDP (high power)")

gridExtra::grid.arrange(f1, f2, ncol =2)
```

@@

## susieI (1)
* Regions: all regions, 500kb uniform regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param1susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

```{r pip1susie, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter1susie, fig.height=4, fig.width=8}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
```

## susieI (2)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param6susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.susieI.txt")

## susieI (3)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=10`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param8susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config8.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config8.susieI.txt")

@@

```{r pip1ser, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter1ser, fig.height=4, fig.width=10}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
```

@@

```{r pip6ser, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter6ser, fig.height=4, fig.width=10}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
```

@@

```{r pip7ser, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter7ser, fig.height=4, fig.width=10}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
```
